\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2016-01-08 paper/paper.Rnw
Cleanup

With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing? In order to use visualisation effectively, we will get the most benefit if the type of graphic display chosen is best suited to the task at hand, be it EDA, MD, or even the presentation of results.
We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within. This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
The lineup protocol is inspired in part by the police lineup \citep{Wickham}. We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people). If the real data plot is selected by an uninvolved observer as being most different from the other plots there is evidence that the structure of that data has led to a significant difference in the plot (That the criminal is sufficiently different from the innocents). 
This constitutes a rejection of the null hypothesis. If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error. 
$p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup. It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hofmann:2014}. 
If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution. The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test. For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$. 
To control our error rate, it is initially obvious that we can simply change $m$, with an increase in null lineups having an inverse reduction in $\alpha$. We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots. A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hoffman:2014} with:

@@

$$~\mathcal{N}\left\(\mu= \left| \begin{array}{c}0 \\0 \end{array} \right|,\hspace{4mm}\Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right|\right\)$$

The variation in the covariance parameter  allows for control in the correlation in the pairs of data. The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$. True data pairs are generated with $\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$. Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$. Further to this, relationships weaker than this point are often of little interest to analysts. Hence we found it unnecessary to test a more full range of correlations. Simulating data does not produce correlations exactly equal to $\rho$, so imulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$. The null pairs often had correlation generated spuriously through the simulation process. This was particularly problematic for generating small sample data, but setting $\beta = max(0.5,100t)$ for both true and null data reduced spurious correlation to manageable levels.
For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications. To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component. This filter creates a trend by penalising both errors in smoothness and in fit. The trend of the series, $\tau_{t}$ is defined by the equation:

>>>>

With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing? In order to use visualisation effectively, we will get the most benefit if the type of graphic display chosen is best suited to the task at hand, be it EDA, MD, or even the presentation of results.
We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within. This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
The lineup protocol is inspired in part by the police lineup \citep{Wickham}. We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people). If the real data plot is selected by an uninvolved observer as being most different from the other plots there is evidence that the structure of that data has led to a significant difference in the plot (That the criminal is sufficiently different from the innocents). 
This constitutes a rejection of the null hypothesis. If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error. 
$p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup. It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hofmann:2014}. 
If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution. The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test. For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$. 
To control our error rate, it is initially obvious that we can simply change $m$, with an increase in null lineups having an inverse reduction in $\alpha$. We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots. A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:

@@

$$e_{t} \sim \mathcal{N} \left( \mu = \left| \begin{array}{c}0 \\ 0 \end{array} \right| , \hspace{4mm} \Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right| \right)$$

The variation in the covariance parameter allows for control in the correlation in the pairs of data. The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$. True data pairs are generated with $\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$. Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$. Further to this, relationships weaker than this point are often of little interest to analysts. Hence we found it unnecessary to test a more full range of correlations. Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$. The null pairs often had correlation generated spuriously through the simulation process. This was particularly problematic for generating small sample data, but setting $\beta = max(0.5,100t)$ for both true and null data reduced spurious correlation to manageable levels.
For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications. To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component. This filter creates a trend by penalising both errors in smoothness and in fit. The trend of the series, $\tau_{t}$ is defined by the equation:
